Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression

### 2016-07-29 manuscript.tex
minor edits
simulation section edits

@@

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1). ). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate. To address this issue, we used a maximum likelihood approach implemented in the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio. Here, [explain more]. Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males. As no significant difference was observed, we report only results from females.

@@

\subsection*{Simulations of Background Selection and Selective Sweeps}

>>>>

%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
%\setlength{\marginparwidth}{1.25cm}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
\newcommand{\X}{\textcolor{red}{\bf X\,}}

@@

\subsection*{Estimating neutral polymorphism on sex chromosomes and autosomes}
For each locus in our analysis, we calculated Watterson’s (1975) estimator of the population parameter $\theta=4N_{e}\mu$, where $N_{e}$ is the effective population size, and $\mu$ is the mutation rate \citep{watterson1975}, using a modified version of the Perl program Polymorphurama \citep{bachtrog2006}. To compare sex-linked and autosomal loci, we calculated the average value of $\theta$, weighted by the number of synonymous sites in each gene (Figure 2; Table 1). ). We obtained 95 percent confidence intervals for our estimates of the X/A and Y/A diversity ratios by bootstrapping by gene using the BCa method \citep{efron1994} implemented in the Boot package in R \citep{canty2012boot}, and calculating X/A and Y/A diversity on each iteration for 20000 replicates each. Bootstrapping was conducted on the final filtered set of 173 sex-linked, and 12355 autosomal genes from the Texas race, and separately for the 176 sex-linked and 11349 autosomal genes from the North Carolina race. Note that the lack of recombination on the Y chromosome implies that assumptions of independence are likely violated, since Y-linked loci are expected to be non-independent, and therefore the true uncertainty in our Y/A diversity estimate obtained by this method is likely an underestimate. To address this issue, we used a maximum likelihood approach using the the software \citep{wright2004hka} to obtain independent estimates of the Y/A ratio (\X Supp. Figures). Here, \X{...explain more}. Finally, we also tested whether estimates of diversity on the X chromosome calculated from females were consistent with estimates from phased X sequences from males. As no significant difference was observed (\X Supp. fig.), we report only results from females.

@@

\subsection*{Simulations of Background Selection and Selective Sweeps}
To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a

\josh{
\textit{to do}
\begin{itemize}
\item \textit{describe parameters used and justify}
\item \textit{do both purifying and positive selection sims}
\end{itemize}
}

We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}. To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data. To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$. We then calculated the average \theta over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure). Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
